---
title: "Visualising numerical data"
subtitle: "<br><br> Data Science in a Box"
author: "[datasciencebox.org](https://datasciencebox.org/)"
output:
  xaringan::moon_reader:
    css: ["https://gavinfay.github.io/fdd-workshops/course-materials/slides/xaringan-themer.css", "https://gavinfay.github.io/fdd-workshops/course-materials/slides/slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "../setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(openintro)
loans_full_schema <- loans_full_schema %>%
  mutate(grade = factor(grade, ordered = TRUE))
loans <- loans_full_schema
```

---

class: middle

# Visualizing numerical data

---

## Describing shapes of numerical distributions

- shape:
    - skewness: right-skewed, left-skewed, symmetric (skew is to the side of the longer tail)
    - modality: unimodal, bimodal, multimodal, uniform
- center: mean (`mean`), median (`median`), mode (not always useful)
- spread: range (`range`), standard deviation (`sd`), inter-quartile range (`IQR`)
- unusual observations

---

class: middle

# Histogram

---

## Histogram

```{r message = TRUE, out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram()
```

---

## Histograms and binwidth

.panelset[
.panel[.panel-name[binwidth = 1000]
```{r out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 1000)
```
]
.panel[.panel-name[binwidth = 5000]
```{r out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000)
```
]
.panel[.panel-name[binwidth = 20000]
```{r out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 20000)
```
]
]

---

## Customizing histograms

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "hist-custom", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r hist-custom, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = loan_amount)) +
  geom_histogram(binwidth = 5000) +
  labs( #<<
    x = "Loan amount ($)", #<<
    y = "Frequency", #<<
    title = "Amounts of Lending Club loans" #<<
  ) #<<
```
]
]

---

## Fill with a categorical variable

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "hist-fill", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r hist-fill, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) + #<<
  geom_histogram(binwidth = 5000,
                 alpha = 0.5) + #<<
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  )
```
]
]

---

## Facet with a categorical variable

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "hist-facet", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r hist-facet, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) + 
  geom_histogram(binwidth = 5000) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  ) +
  facet_wrap(~ homeownership, nrow = 3) #<<
```
]
]

---

class: middle

# Density plot

---

## Density plot

```{r}
ggplot(loans, aes(x = loan_amount)) +
  geom_density()
```

---

## Density plots and adjusting bandwidth

.panelset[
.panel[.panel-name[adjust = 0.5]
```{r out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 0.5)
```
]
.panel[.panel-name[adjust = 1]
```{r out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 1) # default bandwidth
```
]
.panel[.panel-name[adjust = 2]
```{r out.width = "50%"}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2)
```
]
]

---

## Customizing density plots

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "density-custom", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r density-custom, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2) +
  labs( #<<
    x = "Loan amount ($)", #<<
    y = "Density", #<<
    title = "Amounts of Lending Club loans" #<<
  ) #<<
```
]
]

---

## Adding a categorical variable

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "density-cat", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r density-cat, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) + #<<
  geom_density(adjust = 2, 
               alpha = 0.5) + #<<
  labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans", 
    fill = "Homeownership" #<<
  )
```
]
]

---

class: middle

# Box plot

---

## Box plot

```{r}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot()
```

---

## Box plot and outliers

```{r}
ggplot(loans, aes(x = annual_income)) +
  geom_boxplot()
```

---

## Customizing box plots

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "box-custom", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r box-custom, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = NULL,
    title = "Interest rates of Lending Club loans"
  ) +
  theme( #<<
    axis.ticks.y = element_blank(), #<<
    axis.text.y = element_blank() #<<
  ) #<<
```
]
]

---

## Adding a categorical variable

.panelset[
.panel[.panel-name[Plot]
```{r ref.label = "box-cat", echo = FALSE, warning = FALSE}
```
]
.panel[.panel-name[Code]
```{r box-cat, fig.show = "hide", warning = FALSE}
ggplot(loans, aes(x = interest_rate,
                  y = grade)) + #<<
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = "Grade",
    title = "Interest rates of Lending Club loans",
    subtitle = "by grade of loan" #<<
  )
```
]
]

<!-- --- -->

<!-- class: middle -->

<!-- # Relationships numerical variables -->

<!-- --- -->

<!-- ## Scatterplot -->

<!-- ```{r warning = FALSE} -->
<!-- ggplot(loans, aes(x = debt_to_income, y = interest_rate)) + -->
<!--   geom_point() -->
<!-- ``` -->

<!-- --- -->

<!-- ## Hex plot -->

<!-- ```{r warning = FALSE} -->
<!-- ggplot(loans, aes(x = debt_to_income, y = interest_rate)) + -->
<!--   geom_hex() -->
<!-- ``` -->

<!-- --- -->

<!-- ## Hex plot -->

<!-- ```{r warning = FALSE} -->
<!-- ggplot(loans %>% filter(debt_to_income < 100),  -->
<!--        aes(x = debt_to_income, y = interest_rate)) + -->
<!--   geom_hex() -->
<!-- ``` -->
